Exception Handling Bug Hazards in Android

### 2015-10-08 SPRINGER_JOURNAL/msr_journal.tex
coders sentence
including research questions on the mining study - following Christoph's suggestions

The goal of our mining study is to explore to what extent exception stack traces 
contained in issues of Android projects (hosted in GitHub and Google Code), can reveal 
\emph{bug hazards} in the exception-related code of both the applications and the underlying framework. 
As mentioned before, in this context \emph{bug hazards} are the characteristics of exception-related code 
that favor the introduction of failures such as uncaught exceptions. 
To support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Moreover, we use manual inspection to augment
 the understanding of stack traces and support further discussions and insights (Section~\ref{sec:manual}).
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.

@@

This section presents the results of the study; this presentation is centered around 
the information distilled from stack traces: (i) the root exceptions (i.e., the exceptions that caused the stack traces);
 (ii) the exception types (i.e, Checked, Runtime, Error, Throwable) and (iii) the exception wrappings. 
Each piece information is then analyzed in detail to 
check whether they can reveal bug harzards on the exception handling code -
related to (i) specific violations of the best practices presented in Section~\ref{sec:best},
or (ii) the general use of exception handling to suport robust development.

\subsubsection{Can the root exceptions reveal bug harzards?}
After distilling the information available on the exception stack traces, we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such an exception was a method defined on
the Android platform, the Android libcore, the application itself or a third-party library -
 following the classification presented in Table  \ref{tab:signalers}.

@@

\subsubsection{Can the exception types reveal bug hazards?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

\subsubsection{Can the exception wrappings reveal bug hazards?}

@@

When developing Java-based applications it is inevitable to deal with exceptions. Consequently, our first question explores how developers deal with the exception handling code in Android development:

\emph{\textbf{RQ 1:} How often do developers deal with exception handling code while developing Android apps?}

To make the analysis easier, we further refine RQ1 in subquestions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}. Moreover, another question that shows us how developers may deal with the EH code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}

@@

\emph{\textbf{RQ 2:} How do NullPointerExceptions impact the development of robust Android apps?}

\emph{\textbf{RQ 3:} How do cross-type wrappings impact the development of robust Android apps?}

\emph{\textbf{RQ 4:} Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}

%\emph{\textbf{RQ 4:}  Are developers aware of the non-declared checked exceptions?}

@@

\emph{\textbf{RQ 5:} How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}

@@

We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey responses, to answer some of the research questions. The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data.

@@

\bigskip
\noindent\emph{\textbf{Q1:How do developers deal with the exception handling in Android development?}}
\bigskip

@@

\noindent\emph{\textbf{Q 2: How do the NullPointerExceptions impact the development of robust Android apps?}}

@@

\noindent\emph{\textbf{RQ 3: How do Cross-Type Wrappings impact the development of robust Android apps?}}

@@

\bigskip
\noindent\emph{\textbf{RQ 4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}}
\bigskip

@@

\noindent\emph{\textbf{RQ5: How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}}

>>>>

Our mining study was guided by a general research question: \emph{RQ 1: Can the information available in exception stack traces 
 reveal exception handling \emph{bug hazards} in both the Android applications and framework?}
As mentioned before, in this context \emph{bug hazards} are the characteristics of exception-related code 
that favor the introduction of failures such as uncaught exceptions.

Our study focused on the exception stack traces contained in issues of Android projects (hosted in GitHub and Google Code).
And to support our investigation, we developed a tool called ExceptionMiner (Section~\ref{sec:exceptionminer})
which extracts the exception stack traces embedded on issues, 
and combines stack trace information with source code and bytecode 
analysis. Moreover, we use manual inspection to augment
 the understanding of stack traces and support further discussions and insights (Section~\ref{sec:manual}).
In this study we explore the domain quantitatively and highlight interesting cases by 
exploring cases qualitatively.

@@

This section presents the results of our study that was guided by the general research question:
 \emph{RQ 1: Can the information available in exception stack traces reveal exception handling 
bug hazards in both the Android applications and framework?}
To make the analysis easier, we further refine this question in subquestions, each one focusing on pieces of information
 distilled from stack traces, more specifically: (i) the root exceptions (i.e., the exceptions that caused the stack traces);
 (ii) the exception types (i.e, Checked, Runtime, Error, Throwable) and (iii) the exception wrappings. 
Hence, this section is centered around the following subquestions:  \emph{RQ 1.1: Can the root exceptions reveal bug harzards?}; 
 \emph{RQ 1.2 Can the exception types reveal bug hazards?}; and  \emph{RQ 1.3 Can the exception wrappings reveal bug hazards?}.

In this section each piece information is then analyzed in detail to 
check whether they can reveal bug harzards on the exception handling code -
related to (i) specific violations of the best practices presented in Section~\ref{sec:best},
or (ii) the general use of exception handling to suport robust development.

%\subsubsection{RQ 1.1 Can the root exceptions reveal bug harzards?}
\noindent\emph{RQ 1.1 Can the root exceptions reveal bug harzards?}

After distilling the information available on the exception stack traces, we could find 
the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:topten} presents a list of the top 10 root exceptions found in the study - 
 ranked by the number of distinct projects on which they were reported. 
This table also shows how many times the signaler of such an exception was a method defined on
the Android platform, the Android libcore, the application itself or a third-party library -
 following the classification presented in Table  \ref{tab:signalers}.

@@

%\subsubsection{Can the exception types reveal bug hazards?}
\noindent\emph{RQ 1.2 Can the exception types reveal bug hazards?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

%\subsubsection{Can the exception wrappings reveal bug hazards?}
\noindent\emph{RQ 1.3 Can the exception wrappings reveal bug hazards?}

@@

The general research question guided our exploratory survey was the following: \emph{RQ 2: How developers deal with the exception handling code in Android apps and what are the developers' perspectives about the main bug hazards found during the mining study?} This general research question was broken in a set of research questions that are answered by the exploratory survey.

When developing Java-based applications it is inevitable to deal with exceptions. Hence, our first question explores how developers deal with the exception handling code in Android development:

\emph{\textbf{RQ 2.1:} How do developers deal with exception handling code while developing Android apps?}

To make the analysis easier, we further refine this question in subquestions, such as: \emph{How often do developers handle exceptions?}; \emph{How often do developers throw exceptions?}. Moreover, another question that shows us how developers may deal with the EH code is: \emph{Do developers know about Java EH best practices and/or Android specific EH best practices?}

@@

\emph{\textbf{RQ 2.2:} How do NullPointerExceptions impact the development of robust Android apps?}

\emph{\textbf{RQ 2.3:} How do cross-type wrappings impact the development of robust Android apps?}

\emph{\textbf{RQ 2.4:} Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}

%\emph{\textbf{RQ 2.5:}  Are developers aware of the non-declared checked exceptions?}

@@

\emph{\textbf{RQ 2.5:} How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}

@@

We used grounded theory coding~\cite{charmaz2006}, iterating through our open-ended survey responses, to answer some of the research questions. The grounded theory coding used in this study consists of two phases: (1) initial coding entails a close reading of the data and (2) later we use focused coding ~\cite{charmaz2006} to pinpoint and develop the most salient themes~\cite{charmaz2006} in the analyzed data. The coding process was performed by two coders who agreed upon the list of themes that emerged from the answers'set of each question.

@@

\bigskip
\noindent\emph{RQ 2.1: How do developers deal with the exception handling in Android development?}
\bigskip

@@

\noindent\emph{\textbf{RQ 2.2: How do the NullPointerExceptions impact the development of robust Android apps?}}

@@

\noindent\emph{\textbf{RQ 2.3: How do Cross-Type Wrappings impact the development of robust Android apps?}}

@@

\bigskip
\noindent\emph{\textbf{RQ 2.4: Are developers aware of the robustness threats caused by JNI undocumented checked exceptions?}}
\bigskip

@@

\noindent\emph{\textbf{RQ 2.5: How does the exception handling code affect the development of robust apps and how do developers prevent apps from crashing?}}
